---
title: Introduction
tutorial:
  id: introduction
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
    css: ["css/nu-theme.css"]
runtime: shiny_prerendered
description: Tutorial for Introduction
---

<!-- Modification of r4ds.tutorials author: David Kane -->

```{r setup, include = FALSE}

library(learnr)
library(tidyverse)
library(gradethis)
library(sortable)
library(tutorialExtras)

gradethis_setup()

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

```

<!-- ```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")} -->
<!-- ``` -->

<!-- ```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")} -->
<!-- ``` -->

```{r, context = "server"}
grade_server("grade")
```

```{r Name}
question_text("Name:",
              answer_fn(function(value){
                              if(length(value) >= 1 ) {
                                return(mark_as(TRUE))
                                }
                              return(mark_as(FALSE) )
                              }),
              correct = "submitted",
              allow_retry = FALSE )
```

```{r}
grade_button_ui(id = "grade")
```

## Introduction
### 

This tutorial covers the [Introduction](https://r4ds.hadley.nz/intro.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

The tutorials are provide multiple choice questions to check student understanding and code chunks to practice coding along with the associated chapter. This chapter has very little R code. So, in the spirit of getting started, we include some of our own. 

In this tutorial, we will only work within the Exercise Code Blocks. However, you can also put the code commands into the Console tab in the lower left pane of RStudio. 

Each question allows 3 attempts without penalty. After the third attempt a 10% deduction per attempt will occur.

<!-- This tutorial assumes that you have already completed the Getting Started tutorial in the [**tutorial.helpers**](https://ppbds.github.io/tutorial.helpers/) package. If you haven't, do so now. It is quick! -->


## What you will learn
### 

### Exercise 1

```{r what-learn-1}
quiz(
  caption = "",
  question_rank("In a typical data science project what is the typical workflow order?",
           answer(c("tidy your data", "wrangle your data", "visualize your data", "model your data", "communicate your data"), correct = TRUE),
           allow_retry = TRUE,
           #random_answer_order = TRUE,
           options = sortable_options())
)
```


## How this book is organized
### 

The previous description of the tools of data science is organized roughly according to the order in which you use them in an analysis (although, of course, you’ll iterate through them multiple times). In our experience, however, learning data importing and tidying first is suboptimal because, 80% of the time, it’s routine and boring, and the other 20% of the time, it’s weird and frustrating. That’s a bad place to start learning a new subject!

```{r how-organized-1}
quiz(
  caption = "",
  question("What topic will this book start with?",
           answer("visualizing", correct = TRUE),
           answer("tidying"),
           answer("wrangling"),
           answer("communicating"),
           answer("importing"),
           allow_retry = TRUE,
           random_answer_order = TRUE)
)
```

## What you won’t learn

[*R for Data Science (2e)*](https://r4ds.hadley.nz/) and these associated tutorials cover a lot of material. But we can't cover everything. 

### 

```{r not-learn-1}
quiz(
  caption = "",
  question("What topics won't be covered with this book?",
           answer("Python", correct = TRUE),
           answer("Modeling", correct = TRUE),
           answer("Big Data", correct = TRUE),
           answer("Importing"),
           answer("Tidying"),
           answer("Wrangling"),
           allow_retry = TRUE,
           random_answer_order = TRUE)
)
```


## Prerequisites
### 

If you are running this tutorial, then you probably already know about R and RStudio. If, for some reason you don't, this [Getting Started chapter](https://ppbds.github.io/primer/getting-started.html) is the best place to start.

This book will use many packages. Install all the necessary packages by simply copying/pasting this code in the Console of RStudio.

```
install.packages(
  c("arrow", "babynames", "curl", "duckdb", "gapminder", 
    "ggrepel", "ggridges", "ggthemes", "hexbin", "janitor", "Lahman", 
    "leaflet", "maps", "nycflights13", "openxlsx", "palmerpenguins", 
    "repurrrsive", "tidymodels", "writexl")
  )
```

### Exercise 1

Before you start doing data science, you must load the packages you are going to use. Use the function `library()` to load the **tidyverse** package.

```{r prerequisites-1, exercise = TRUE}

```

```{r prerequisites-1-hint-1, eval = FALSE}
library(tidyverse)
```

<!-- ```{r prerequisites-1-test, include = FALSE} -->
<!-- library(tidyverse) -->
<!-- ``` -->

```{r prerequisites-1-solution, exercise.reveal_solution = FALSE}
library(tidyverse)
```

```{r prerequisites-1-check}
grade_this_code()
```

### 

Nothing is returned, which is often the case with R code. You can "Run code" as many times as you want to test your answer but make sure you click "Submit Answer" to record your submission.

### 

Note that "library" and "package" mean the same things in R. We have different words for historical reasons. However, only the `library()` command will load a package/library, giving us access to the functions and data which it contains.

### Exercise 2

If you are ever stuck while coding, R has *help pages*.

Type `?diamonds` to look up the help page for the `diamonds` tibble from the **ggplot2** package, which is one of the core packages in the *Tidyverse*.

```{r prerequisites-2a, exercise = TRUE}

```

```{r prerequisites-2a-hint-1, eval = FALSE}
?diamonds
```


```{r prerequisites-2a-solution, exercise.reveal_solution = FALSE}
?diamonds
```

```{r prerequisites-2a-check}
grade_this_code()
```

Copy/paste the Description from the help page into the box below.

```{r prerequisites-2b}
question_text(NULL,
	answer("A dataset containing the prices and other attributes of almost 54,000 diamonds.", correct = TRUE),
	answer("A dataset containing the prices and other attributes of almost 54,000 diamonds. The variables are as follows:", correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

You can find help about an entire package with `help(package = "ggplot2")`. It is confusing, but unavoidable, that package names are sometimes unquoted, as in `library(ggplot2)`, and sometimes quoted, as in `help(package = "ggplot2")`. If one does not work, try the other.

## Running R code
###

### Exercise 1

Assign the value of `sqrt(144)` to the variable `x`. Remember to use the assignment operator `<-`.

```{r running-code-1, exercise = TRUE}

```

```{r running-code-1-hint-1, eval = FALSE}
variable_name <- sqrt(...)
```

<!-- ```{r running-code-1-test, include = FALSE} -->
<!-- x <- sqrt(144) -->
<!-- ``` -->

```{r running-code-1-solution, exercise.reveal_solution = FALSE}
x <- sqrt(144)
```

```{r running-code-1-check}
grade_this_code()
```


### Exercise 2
### 

Data frames, also referred to as "tibbles", are spreadsheet-type data sets. 

### 

In data science, it's always a good idea to explore the data before you use it. You can simply examine a data set by typing its name. Let's examine the contents of `diamonds` data set. Type `diamonds` in the line below.

```{r running-code-2a, exercise = TRUE}

```

```{r running-code-2a-hint-1, eval = FALSE}
diamonds
```

<!-- ```{r running-code-2a-test, include = FALSE} -->
<!-- diamonds -->
<!-- ``` -->

```{r running-code-2a-solution, exercise.reveal_solution = FALSE}
diamonds
```

```{r running-code-2a-check}
grade_this_code()
```

### 

When we type the name of a tibble in the R Console, we are implicitly using the `print()` function. Try it by using `print()` on `diamonds`.

```{r running-code-2b, exercise = TRUE}

```

```{r running-code-2b-hint-1, eval = FALSE}
print(...)
```

<!-- ```{r running-code-2b-test, include = FALSE} -->
<!-- print(diamonds) -->
<!-- ``` -->

```{r running-code-2b-solution, exercise.reveal_solution = FALSE}
print(diamonds)
```

```{r running-code-2b-check}
grade_this_code()
```

This produces the same result as just typing `diamonds`. But `print()` is more flexible, allowing us to specify different numbers of rows and columns. You can use the `width` argument to change how many columns appear. `print(diamonds, width = Inf)` will cause all columns to be printed.

###

Print out the first three rows of `diamonds`. Note that the number of rows printed is controlled by the `n` argument.

```{r running-code-2c, exercise = TRUE}

```

```{r running-code-2c-hint-1, eval = FALSE}
print(dataset, n = ?)
```

<!-- ```{r running-code-2c-test, include = FALSE} -->
<!-- print(diamonds, n = 3) -->
<!-- ``` -->

```{r running-code-2c-solution, exercise.reveal_solution = FALSE}
print(diamonds, n = 3)
```

```{r running-code-2c-check}
grade_this_code()
```


### 

Whenever we print a tibble, the number of rows and columns is displayed at the top:

````
A tibble: rows × columns
````

You can also see the variable type under each of the column names.

```{r running-code-2d}
quiz(
  caption = "",
  question_numeric("How many observations are in the diamonds dataset?",
           answer(53940, correct = TRUE),
           step = 1,
           allow_retry = TRUE),
  question_numeric("How many variables are in the diamonds dataset?",
           answer(10, correct = TRUE),
           step = 1,
           allow_retry = TRUE),
  question_numeric("How many of the variables are numeric?",
           answer(7, correct = TRUE),
           step = 1,
           allow_retry = TRUE)
)
```


### Exercise 3

Now use `glimpse()` to examine the `diamonds` data set. To learn more about the `glimpse()` function, run `?glimpse`.

```{r running-code-3, exercise = TRUE}

```

```{r running-code-3-hint-1, eval = FALSE}
___(dataset)
```

<!-- ```{r running-code-3-test, include = FALSE} -->
<!-- glimpse(diamonds) -->
<!-- ``` -->

```{r running-code-3-solution, exercise.reveal_solution = FALSE}
glimpse(diamonds)
```

```{r running-code-3-check}
grade_this_code()
```

### 

`glimpse()` displays columns running down the page and the data running across across. Note how the "type" of each variable is listed next to the variable name. For example, `price` is listed as `<int>`, meaning that it is an integer variable.

`view()` is another useful function, but, because it is interactive, we should not use it within a tutorial. 

### Exercise 4

Run `summary()` on `diamonds`.

```{r running-code-4, exercise = TRUE}

```

```{r running-code-4-hint-1, eval = FALSE}
summary(...)
```

<!-- ```{r running-code-4-test, include = FALSE} -->
<!-- summary(diamonds) -->
<!-- ``` -->

```{r running-code-4-solution, exercise.reveal_solution = FALSE}
summary(diamonds)
```

```{r running-code-4-check}
grade_this_code()
```


### Exercise 5

*Code comments* are text placed after a `#` symbol. Nothing will be run after a `#` symbol, which is useful if you want to write human readable comments in your code.

### 

Press "Run Code." Afterwards, add the `#` and re-run the chunk. You should no longer see a result. Submit your commented out code.

```{r running-code-5, exercise = TRUE}
sqrt(144)
```


```{r running-code-5-solution, exercise.reveal_solution = FALSE}
# sqrt(144)
```

```{r running-code-5-check}
grade_this_code()
```


### Exercise 6

The double colon --- `::` --- notation allows us to access the inside of a package even if we have not loaded it.

The `clean_names()` function is from the `janitor` package and is used to convert variable names to only include the `_` character, numbers, and letters.

###

Use the double colon to run the `clean_names()` function on the `diamonds` dataset.

```{r running-code-6, exercise = TRUE}

```

```{r running-code-6-hint-1, eval = FALSE}
package::function_name(...)
```

<!-- ```{r running-code-6-test, include = FALSE} -->
<!-- janitor::clean_names(diamonds) -->
<!-- ``` -->

```{r running-code-6-solution, exercise.reveal_solution = FALSE}
janitor::clean_names(diamonds)
```

```{r running-code-6-check}
grade_this_code()
```


### Exercise 7

The `+` sign is used to "chain" different pieces of plot creation code together. When doing data analysis, we use the "pipe" symbol --- `|>` --- to do the same thing between different pieces of code which manipulate the data.

###

As a simple example, re-do the previous problem but this time "pipe" the `diamonds` tibble into the function.  

```{r running-code-7, exercise = TRUE}

```

```{r running-code-7-hint-1, eval = FALSE}
data |> 
  package::function_name()
```

<!-- ```{r running-code-7-test, include = FALSE} -->
<!-- diamonds |>  -->
<!--   janitor::clean_names() -->
<!-- ``` -->

```{r running-code-7-solution, exercise.reveal_solution = FALSE}
diamonds |> 
  janitor::clean_names()
```

```{r running-code-7-check}
grade_this_code()
```

### 

Note the language. We write "pipe this to that." That is, we pipe the `diamonds` tibble to the `clean_names()` command. This accomplishes the same effect as simply running `clean_names(diamonds)`, but allows us to string together several commands in a row.

## Summary

This tutorial covered the [Introduction](https://r4ds.hadley.nz/intro.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

###

Throughout these tutorials, we use a consistent set of conventions to refer to code:

Functions are displayed in a code font and followed by parentheses, like `sum()` or `mean()`.

Other R objects (such as data or function arguments) are in a code font, without parentheses, like `flights` or `x`.

Sometimes, to make it clear which package an object comes from, we’ll use the package name followed by two colons, like `dplyr::mutate()` or `nycflights13::flights`. This is also valid R code.


<!-- ```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")} -->
<!-- ``` -->

###

Once you are finished:

-   Click the 'Download Grade' button below. This will download an html document of your grade summary.
-   Submit the downloaded html to your learning management system.

```{r}
grade_print_ui("grade")
```