---
title: Layers
tutorial:
  id: layers
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
    css: ["css/nu-theme.css"]
runtime: shiny_prerendered
description: 'Tutorial for Chapter 9: Layers'
---

<!-- Modification of r4ds.tutorials author: David Kane -->

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(ggridges)
library(maps)
library(gradethis)
library(tutorialExtras)

gradethis_setup()

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

nz <- map_data("nz")

polling_data <- read_csv("data/vis4_polling_data.csv",
              col_types = cols(createddate = col_date(),
                               approval_type = col_character(),
                               rate = col_double()))
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```


```{r, context = "server"}
grade_server("grade")
```

```{r Name}
question_text("Name:",
              answer_fn(function(value){
                              if(length(value) >= 1 ) {
                                return(mark_as(TRUE))
                                }
                              return(mark_as(FALSE) )
                              }),
              correct = "submitted",
              allow_retry = FALSE )
```

```{r}
grade_button_ui(id = "grade")
```


<!-- Make the plots cooler! Add links to  https://ggplot2-book.org/. And https://socviz.co/. Or maybe put these in case studies. Love this one: https://socviz.co/groupfacettx.html#histograms. -->

<!-- DK: The final four plot exercies, which are not part of R4DS, need to be dealt with. Maybe turn these into a graduation exercise with one repo, one QMD which contains the three plots, then publish them? Some of these plots are not very good. Maybe move some of them to other tutorials? -->

## Introduction
### 

This tutorial covers [Chapter 9: Layers](https://r4ds.hadley.nz/layers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to use [**ggplot2**](https://ggplot2.tidyverse.org/) and related packages to create persuasive graphics. Key functions include [`geom_smooth()`](https://ggplot2.tidyverse.org/reference/geom_smooth.html),  [`geom_histogram()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html), and  [`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html).


## Aesthetic mappings
### 

We’ll start with a deeper dive into aesthetic mappings. 

### Exercise 1

Load up the **tidyverse** library.

```{r aesthetic-mappings-1, exercise = TRUE}

```

```{r aesthetic-mappings-1-hint-1, eval = FALSE}
library(...)
```

```{r aesthetic-mappings-1-solution, exercise.reveal_solution = FALSE}
library(tidyverse)
```

```{r aesthetic-mappings-1-check}
grade_this_code()
```

### 

> “The greatest value of a picture is when it forces us to notice what we never expected to see.” — John Tukey

### Exercise 2

Run `glimpse()` on `mpg`.

```{r aesthetic-mappings-2, exercise = TRUE}

```

```{r aesthetic-mappings-2-hint-1, eval = FALSE}
glimpse(...)
```

```{r aesthetic-mappings-2-solution, exercise.reveal_solution = FALSE}
glimpse(mpg)
```

```{r aesthetic-mappings-2-check}
grade_this_code()
```

### 

The `mpg` tibble bundled with the **ggplot2** package contains 234 observations on 38 car models. The key variables for us include:

* **displ**: A car’s engine size, in liters. A numerical variable.

* **hwy**: A car’s fuel efficiency on the highway, in miles per gallon (mpg). A car with a low fuel efficiency consumes more fuel than a car with a high fuel efficiency when they travel the same distance. A numerical variable.

* **class**: Type of car. A categorical variable.


### Exercise 3

Run `ggplot()`.

- Set the first argument, `data`, equal to `mpg`.

- Set the second argument, `mapping`, equal to the `aes()` function, setting the `x` argument to `displ` and the `y` argument to equal `hwy`.  

- Add `geom_point()`

```{r aesthetic-mappings-3, exercise = TRUE}

```

```{r aesthetic-mappings-3-hint-1, eval = FALSE}
ggplot(data = ..., mapping = ...) +
  ...()
```

```{r aesthetic-mappings-3-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  geom_point()
```

```{r aesthetic-mappings-3-check}
grade_this_code()
```

### 

A scatter plot like this is the most common way to look at data. If there is a lot of overlap among the data points, you may want to use [`geom_jitter()`](https://ggplot2.tidyverse.org/reference/geom_jitter.html) to separate the points.


### Exercise 4

Set the `color` argument within `aes()` to `class`. Each point will have a color based on the `class` of the car.

```{r aesthetic-mappings-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-4-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = ...)) +
  geom_point()
```

```{r aesthetic-mappings-4-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy,
                color = class)) +
  geom_point()
```

```{r aesthetic-mappings-4-check}
grade_this_code()
```

### 

The plot is much richer, even though it isn't any larger, because we have used color to incorporate information about the type of car. Notice how the SUVs cluster in the lower right. Their engines are (mostly) bigger, and they all get worse gas mileage.

### Exercise 5

Replace the `color` argument with the `shape` argument within `aes()`. Set the value to `class`, as before. Each point will have a shape based on the `class` of the car.

```{r aesthetic-mappings-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-5-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                shape = ...)) +
  geom_point()
```

```{r aesthetic-mappings-5-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy,
                shape = class)) +
  geom_point()
```

```{r aesthetic-mappings-5-check}
grade_this_code()
```

When class is mapped to shape, we get two warnings:

````
1: The shape palette can deal with a maximum of 6 discrete values because more than 6 becomes difficult to discriminate; you have 7. Consider specifying shapes manually if you must have them.

2: Removed 62 rows containing missing values (geom_point()).
````
 
### 

Since `ggplot()` will only use six shapes at a time, by default, additional groups will go unplotted when you use the `shape` aesthetic. The second warning is related – there are 62 SUVs in the dataset and they’re not plotted.

### Exercise 6

Replace the `shape` argument with the `size` argument within `aes()`. Set the value to `class`, as before. Each point will have a size based on the `class` of the car.

```{r aesthetic-mappings-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-6-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                size = ...)) +
  geom_point()
```

```{r aesthetic-mappings-6-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy,
                size = class)) +
  geom_point()
```

```{r aesthetic-mappings-6-check}
grade_this_code()
```

You should see a warning message:

````
Using size for a discrete variable is not advised. 
````

### 

Mapping an unordered discrete (categorical) variable, like `class`, to an ordered aesthetic, like `size` is generally not a good idea because it implies a ranking that does not in fact exist, hence the warning.


### Exercise 7

Replace the `size` argument with the `alpha` argument within `aes()`. Set the value to `class`, as before. Each point will have a transparency based on the `class` of the car.

```{r aesthetic-mappings-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-7-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                alpha = ...)) +
  geom_point()
```

```{r aesthetic-mappings-7-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy,
                alpha = class)) +
  geom_point()
```

```{r aesthetic-mappings-7-check}
grade_this_code()
```

You should see a warning message, as above:

````
Using alpha for a discrete variable is not advised. 
````

### 

Once you map an aesthetic, **ggplot2** takes care of the rest. It selects a reasonable scale to use with the aesthetic, and it constructs a legend that explains the mapping between levels and values. For x and y aesthetics, **ggplot2** does not create a legend, but it creates an axis line with tick marks and a label. The axis line provides the same information as a legend; it explains the mapping between locations and values.


### Exercise 8

You can also set the visual properties of your geom manually as an argument of your geom function (outside of `aes()`) instead of relying on a variable mapping to determine the appearance. Remove the `alpha` argument from your code and add `color = "blue"` **within** the call to `geom_point()`.

```{r aesthetic-mappings-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-8-hint-1, eval = FALSE}
ggplot(mpg, aes(x = ..., 
                y = hwy)) + 
  geom_point(color = "...")
```

```{r aesthetic-mappings-8-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point(color = "blue")
```

```{r aesthetic-mappings-8-check}
grade_this_code()
```

### 

Here, the color doesn’t convey information about a variable, but only changes the appearance of the plot. You’ll need to pick a value that makes sense for that aesthetic:

* The name of a color as a character string, e.g., color = "blue".
* The size of a point in mm, e.g., size = 1.
* The shape of a point as a number, e.g, shape = 1. 

## Geometric objects
### 

So far we have discussed aesthetics that we can map or set in a scatterplot, when using a point geom. You can learn more about all possible aesthetic mappings in the aesthetic specifications [vignette](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html).

### Exercise 1

To change the geom in your plot, change the geom function that you add to `ggplot()`. 

Run `ggplot()`

- Set the first argument equal to `mpg`.

- Set the second argument equal to the `aes()` function, setting the `x` argument to `displ` and the `y` argument to equal `hwy`.  

- This time add `geom_smooth()`.

```{r geometric-objects-1, exercise = TRUE}

```

```{r geometric-objects-1-hint-1, eval = FALSE}
ggplot(..., aes(x = displ, 
                y = hwy)) + 
  geom_smooth()
```

```{r geometric-objects-1-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  geom_smooth()
```

```{r geometric-objects-1-check}
grade_this_code()
```

### 

<!-- Every geom function in **ggplot2** takes a mapping argument, either defined locally in the geom layer or globally in the `ggplot()` layer.  -->

<!-- ### -->

This plot contain the same `x` variable, the same `y` variable, and describes the same data as our scatterplot in the last section. But the plots are not identical. Each plot uses a different geometric object, geom, to represent the data.

### 

Not every aesthetic works with every geom. You could set the shape of a point, but you couldn’t set the “shape” of a line. If you try, **ggplot2** will silently ignore that aesthetic mapping. 

### Exercise 2

Note the message which gets displayed when calling `geom_smooth()`.

````
`geom_smooth()` using method = 'loess' and formula = 'y ~ x'
````

**ggplot2** thinks that `method = 'loess'` and `formula = 'y ~ x'` are good default values, but it wants to be sure that you are aware of them. 

###

First, add `method = 'loess'` to the `geom_smooth()` call.

Second, add `formula = 'y ~ x'`. 

Third, add `se = FALSE`.

```{r geometric-objects-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-2-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_smooth(... = 'loess', 
              ... = 'y ~ x',
              se = ...)
```

```{r geometric-objects-2-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE)
```

```{r geometric-objects-2-check}
grade_this_code()
```

### 

By explicitly defining the `method` and `formula` in your code the message goes away, which you should do.

### 

The `se`, which stands for "standard error," determines whether or not error bars around the line are included. By default, the value of `se` is `TRUE`. We set it to `FALSE` here in order to make the graphic less cluttered.


### Exercise 3

You can set the `linetype` of a line. `geom_smooth()` will draw a different line, with a different `linetype`, for each unique value of the variable that you map to `linetype`. Within the call to `aes()`, add `linetype = drv`.

```{r geometric-objects-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-3-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                linetype = ...)) + 
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE)
```

```{r geometric-objects-3-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                linetype = drv)) + 
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE)
```

```{r geometric-objects-3-check}
grade_this_code()
```

### 

Here, `geom_smooth()` separates the cars into three lines based on their `drv` value, which describes a car’s drive train. One line describes all of the points that have a `4` value. One line describes all of the points that have an `f` value. And one line describes all of the points that have an `r` value. Here, `4` stands for four-wheel drive, `f` for front-wheel drive, and `r` for rear-wheel drive.

### Exercise 4

If this sounds strange, we can make it clearer by overlaying the lines on top of the raw data. To do this, simply add a call to `geom_point()` in between `ggplot()` and `geom_smooth()`. Don't forget to separate different lines with `+`.

Also set `color = drv` within `aes()`.

```{r geometric-objects-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-4-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                linetype = drv,
                color = ...)) + 
  ...() +
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE)
```

```{r geometric-objects-4-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                linetype = drv,
                color = drv)) + 
  geom_point() +
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE)
```

```{r geometric-objects-4-check}
grade_this_code()
```

### 

Notice that this plot contains two geoms in the same graph.

### Exercise 5

The `aes()` values cascade across all the geom's, unless and until we change them ourselves. 

Notice that `geom_point()` simply ignores `linetype` because that is not a valid aesthetic for a point.

It would be more appropriate to define `linetype` locally within the `geom_smooth()` call (instead of globally).

Set `linetype = drv` within the `aes()` of the call to `geom_smooth()`.

```{r geometric-objects-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-5-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = drv)) + 
  ... +
  geom_smooth(...(linetype = ...))
```

```{r geometric-objects-5-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = drv)) + 
  geom_point() +
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE,
              aes(linetype = drv))
```

```{r geometric-objects-5-check}
grade_this_code()
```

### 

Many geoms, like `geom_smooth()`, use a single geometric object to display multiple rows of data. For these geoms, you can set the `group` aesthetic to a categorical variable to draw multiple objects. 

### Exercise 6

The explore the use of the `group` aesthetic, begin with our standard plot, which uses just `aes(x = displ, y = hwy)` in the call to `ggplot()` and just `method = 'loess'`,  `formula = 'y ~ x'`, and `se = FALSE` in the call to `geom_smooth()`.

```{r geometric-objects-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-6-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  geom_smooth(method = ..., 
              ... = 'y ~ x',
              se = ...)
```

```{r geometric-objects-6-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE)
```

```{r geometric-objects-6-check}
grade_this_code()
```

### 

Note how the line extends across the full range of `displ` because we are not splitting up the geom for different values of `drv`.

### Exercise 7

**ggplot2** will draw a separate object for each unique value of the grouping variable. In practice, **ggplot2** will automatically group the data for these geoms whenever you map an aesthetic to a discrete variable (as in the `linetype` example).

Add `aes(group = drv)` to the call to `geom_smooth()`.

```{r geometric-objects-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-7-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE,
              ...(group = ...))
```

```{r geometric-objects-7-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE,
              aes(group = drv))
```

```{r geometric-objects-7-check}
grade_this_code()
```
### 

It is convenient to rely on this feature because the `group` aesthetic by itself does not add a legend or distinguishing features to the geoms.

### Exercise 8

Switch `group = drv` to `color = drv` and add `show.legend = FALSE` to the call to `geom_smooth()`.

```{r geometric-objects-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-8-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE,
              aes(color = drv), 
              ... = FALSE)
```

```{r geometric-objects-8-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(method = 'loess', 
              formula = 'y ~ x',
              se = FALSE,
              aes(color = drv), 
              show.legend = FALSE)
```

```{r geometric-objects-8-check}
grade_this_code()
```

### 

It is sometimes difficult to know if the `color` aesthetic should be set in the `ggplot()` call or in the call to the individual geom. Experiment!


### Exercise 9

To explore some of these concepts, let's construct this plot:

```{r}
geom_p <- ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    color = "red"
  ) +
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    shape = "circle open", size = 3, color = "red"
  )

geom_p
```

Begin with `ggplot(mpg, aes(x = displ, y = hwy))` and use `geom_point()`. 


```{r geometric-objects-9, exercise = TRUE}

```

```{r geometric-objects-9-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, y = ...)) + 
  ...()
```

```{r geometric-objects-9-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```

```{r geometric-objects-9-check}
grade_this_code()
```
### 

The easiest way to build a complex plot is *interactively*. Add one line. See how it looks. Add another line. And so on.


### Exercise 10

Add a second call to `geom_point()`. In that call, set `data` equal to `mpg |> filter(class == "2seater")` and `color` equal to `"red"`.

```{r geometric-objects-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-10-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_point(
    ... = mpg |> filter(class == "2seater"), 
    color = ...
  )
```

```{r geometric-objects-10-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    color = "red"
  )
```

```{r geometric-objects-10-check}
grade_this_code()
```
### 

By default, calls to geoms use whatever values for the `data` argument have been given previously. But, the `data` argument can also be changed. In this case, the second call to `geom_point()` only applies to the five rows for which `class` is "2seater".

### Exercise 11

Add a third call to `geom_point()`, setting `data` equal to `mpg |> filter(class == "2seater")`, `color` equal to `"red"` (as in the previous call), `shape` equal to `"circle open"`, and `size` equal to `3`.

```{r geometric-objects-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-11-hint-1, eval = FALSE}

```

```{r geometric-objects-11-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_point(
    data = mpg |> filter(class == "2seater"),
    color = "red"
  ) +
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    color = "red",
    shape = "circle open",
    size = 3 
  )
```

```{r geometric-objects-11-check}
grade_this_code()
```
### 

We need the two extra calls to `geom_point()` because the first one makes the central point red while the second makes the red circle around it.

### Exercise 12

Geoms are the fundamental building blocks of ggplot2. You can completely transform the look of your plot by changing its geom, and different geoms can reveal different features of your data. 

Make a new plot by hitting "Submit Code".

```{r geometric-objects-12, exercise = TRUE}
ggplot(mpg, aes(x = hwy)) +
  geom_histogram(binwidth = 2)
```

```{r geometric-objects-12-hint-1, eval = FALSE}
ggplot(mpg, aes(x = hwy)) +
  geom_histogram(binwidth = 2)
```

```{r geometric-objects-12-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = hwy)) +
  geom_histogram(binwidth = 2)
```

```{r geometric-objects-12-check}
grade_this_code()
```
### 

The histogram reveals that the distribution of highway mileage is bimodal and right skewed. You can see a similar plot if you replace `geom_histogram(binwidth = 2)` with `geom_density()`.

### Exercise 13

Replace `geom_histogram(binwidth = 2)` with `geom_boxplot()` to create a boxplot with the same data.

```{r geometric-objects-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-13-hint-1, eval = FALSE}
ggplot(mpg, aes(x = hwy)) +
  geom_boxplot()
```

```{r geometric-objects-13-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = hwy)) +
  geom_boxplot()
```

```{r geometric-objects-13-check}
grade_this_code()
```
### 

`geom_boxplot()` helps to highlight the two outlier data points. **ggplot2** provides more than 40 geoms but these don’t cover all possible plots one could make. If you need a different geom, look at extension packages first to see if someone else has already implemented it (see https://exts.ggplot2.tidyverse.org/gallery/ for a sampling).

### Exercise 14

The [**ggridges** package](https://wilkelab.org/ggridges) is useful for making ridgeline plots, which can be useful for visualizing the density of a numerical variable for different levels of a categorical variable. Hit "Submit Code" to see an example.

```{r geometric-objects-14, exercise = TRUE}
library(ggridges)

ggplot(mpg, aes(x = hwy, y = drv, fill = drv, color = drv)) +
  geom_density_ridges(alpha = 0.5, show.legend = FALSE)
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-14-hint-1, eval = FALSE}
ggplot(mpg, aes(x = hwy, y = drv, fill = drv, color = drv)) +
  geom_density_ridges(alpha = 0.5, show.legend = FALSE)
```

```{r geometric-objects-14-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = hwy, y = drv, fill = drv, color = drv)) +
  geom_density_ridges(alpha = 0.5, show.legend = FALSE)
```

```{r geometric-objects-14-check}
grade_this_code()
```
### 

The best place to get a comprehensive overview of all of the geoms **ggplot2** offers, as well as all functions in the package, is the [reference page](https://ggplot2.tidyverse.org/reference). To learn more about any single geom, use the help (e.g., `?geom_smooth`).


## Facets
### 

Facets, created with functions like `facet_wrap()` and `facet_grid()`, are the easiest way to create plot "multiples," copies of the same plot for each value of a third variable.

In Chapter 1 you learned that `facet_wrap()` is used to facet your plot by a single variable. The first argument of `facet_wrap()` is a formula, which you create with `~` followed by a variable name. The variable that you pass to `facet_wrap()` should be categorical, i.e., either character or factor.

### Exercise 1

Start with our usual basic scatterplot, `displ` versus `hwy` from the `mpg` tibble.

Add `facet_wrap(~cyl)` to the end of your scatterplot code. Don't forget the `+` after `geom_point()`.

```{r facets-1, exercise = TRUE}

```

```{r facets-1-hint-1, eval = FALSE}
ggplot(mpg, aes(... = displ, 
                y = ...)) + 
  ...() + 
  facet_wrap(~ ...)
```

```{r facets-1-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point() + 
  facet_wrap(~cyl)
```

```{r facets-1-check}
grade_this_code()
```

### 

[`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than `facet_grid()` because most displays are roughly rectangular.

### Exercise 2

To facet your plot with the combination of two variables, switch from `facet_wrap()` to `facet_grid()`. The first argument of `facet_grid()` is also a formula, but now it’s a double sided formula: rows ~ cols. 

Replace `facet_wrap(~cyl)` with `facet_grid(drv ~ cyl)`.

```{r facets-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-2-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point() + 
  facet_...(... ~ cyl)
```

```{r facets-2-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point() + 
  facet_grid(drv ~ cyl)
```

```{r facets-2-check}
grade_this_code()
```

### 

By default each of the facets share the same scale and range for x and y axes. This is useful when you want to compare data across facets but it can be limiting when you want to visualize the relationship within each facet better. 

### Exercise 3

Setting the scales argument in a faceting function to "free" will allow for different axis scales across both rows and columns, "free_x" will allow for different scales across rows, and "free_y" will allow for different scales across columns.

Add `scales = "free_y"` to the call to `facet_grid()`. Don't forget the comma after `drv ~ cyl`.

```{r facets-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-3-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  facet_grid(drv ~ cyl, ... = ...)
```

```{r facets-3-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  facet_grid(drv ~ cyl, scales = "free_y")
```

```{r facets-3-check}
grade_this_code()
```
### 

[`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) and 
[`facet_grid()`](https://ggplot2.tidyverse.org/reference/facet_grid.html) include lots of useful arguments, but note that the names of those arguments are not always consistent, for historical reasons. Consider `nrow` from `facet_wrap()` versus `rows` from `facet_grid()`.

## Statistical transformations
### 

Instead of simply plotting data, it is sometimes useful to calculate a statistic based on the data and then plot the value of that statistic.

### Exercise 1

The `diamonds` dataset is part of the **ggplot2** package. 

The `diamonds` dataset contains information on about 54,000 diamonds, including the `price`, `carat`, `color`, `clarity`, and `cut` of each diamond. 

###

1. Pipe `diamonds` into `ggplot()`. (Recall, instead of passing in `diamonds` as the first argument to `ggplot()`, we can "pipe" it in with the `|>` symbol.)

2. Add `aes(x = cut)` to the class to `ggplot()`.

3. Add `geom_bar()` as the geom. Don't forget the `+` after the call to `ggplot()`.

```{r statistical-transformations-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-1-hint-1, eval = FALSE}
diamonds |> 
  ...(aes(...)) + 
  ...()
```

```{r statistical-transformations-1-solution, exercise.reveal_solution = FALSE}
diamonds |> 
  ggplot(aes(x = cut)) + 
  geom_bar()
```

```{r statistical-transformations-1-check}
grade_this_code()
```

### 

On the x-axis, the chart displays cut, a variable from diamonds. On the y-axis, it displays count, but count is not a variable in diamonds! It is a "statistical transformation" which **ggplot2** has calculated for us.

### Exercise 2

Pipe `diamonds` into `count(cut)`.

```{r statistical-transformations-2, exercise = TRUE}

```

```{r statistical-transformations-2-hint-1, eval = FALSE}

```

```{r statistical-transformations-2-solution, exercise.reveal_solution = FALSE}
diamonds |>
  count(cut)
```

```{r statistical-transformations-2-check}
grade_this_code()
```
### 

Many graphs, like scatterplots, plot the raw values of your dataset. Other graphs, like bar charts, calculate new values to plot:

* Bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin.

* Smoothers fit a model to your data and then plot predictions from the model.

* Boxplots compute the five-number summary of the distribution and then display that summary as a specially formatted box.

### Exercise 3

`count()` outputs a tibble with values of the variable(s) you selected, along with the number of observations corresponding to each value. We can pipe this output directly into `ggplot()`, along with `aes(x = cut, y = n)`.

```{r statistical-transformations-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-3-hint-1, eval = FALSE}
diamonds |>
  count(cut) |>
  ggplot(aes(x = ..., ... = n))
```

```{r statistical-transformations-3-solution, exercise.reveal_solution = FALSE}
diamonds |>
  count(cut) |>
  ggplot(aes(x = cut, y = n))
```

```{r statistical-transformations-3-check}
grade_this_code()
```
### 

The algorithm used to calculate new values for a graph is called a "stat," short for statistical transformation. In this case, we are calculating the stat by hand, and then using the calculated value `n` to make a plot.

### Exercise 4

Add `geom_bar(stat = 'identity')` to the pipeline.

```{r statistical-transformations-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-4-hint-1, eval = FALSE}

```

```{r statistical-transformations-4-solution, exercise.reveal_solution = FALSE}
diamonds |>
  count(cut) |>
  ggplot(aes(x = cut, y = n)) +
  geom_bar(stat = 'identity')
```

```{r statistical-transformations-4-check}
grade_this_code()
```
### 

The result is the same as the much-more-straight-forward usage of `geom_bar()` which we created above.

### Exercise 5

You might want to override the default mapping from transformed variables to aesthetics. For example, you might want to display a bar chart of proportions, rather than counts.

To do this, set `y = after_stat(prop)` and `group = 1`.

```{r statistical-transformations-5, exercise = TRUE}
ggplot(diamonds, aes(x = cut, y = ___, group = ___)) + 
  geom_bar()
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-5-hint-1, eval = FALSE}
ggplot(diamonds, aes(x = cut, y = after_stat(prop), group = 1)) + 
  geom_bar()
```

```{r statistical-transformations-5-solution, exercise.reveal_solution = FALSE}
ggplot(diamonds, aes(x = cut, y = after_stat(prop), group = 1)) + 
  geom_bar()
```

```{r statistical-transformations-5-check}
grade_this_code()
```
### 

To find the possible variables that can be computed by the stat, look for the section titled “Computed variables” in the [help](https://ggplot2.tidyverse.org/reference/geom_bar.html) for `geom_bar()`.

###

Setting `group`, is used to override the default grouping of `cut`. If we did not set this argument it would instead set each proportion to 100% because the proportion of Fair in Fair is 100%, etc.

### Exercise 6

You might want to draw greater attention to the statistical transformation in your code. For example, you might use stat_summary(), which summarizes the y values for each unique x value, to draw attention to the summary that you’re computing. Hit "Submit Code".

```{r statistical-transformations-6, exercise = TRUE}
ggplot(diamonds) + 
  stat_summary(
    aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-6-hint-1, eval = FALSE}
ggplot(diamonds) + 
  stat_summary(
    aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

```{r statistical-transformations-6-solution, exercise.reveal_solution = FALSE}
ggplot(diamonds) + 
  stat_summary(
    aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

```{r statistical-transformations-6-check}
grade_this_code()
```
### 

**ggplot2** provides more than 20 stats for you to use. Each stat is a function, so you can get help in the usual way, e.g., `?stat_bin`.

## Position adjustments
### 

There’s one more piece of magic associated with bar charts. You can color a bar chart using either the `color` aesthetic, or, more usefully, the `fill` aesthetic.

### Exercise 1

Create a plot which starts with `ggplot(mpg, aes(x = drv, fill = drv))` and then adds, after a `+`, `geom_bar()`.


```{r position-adjustments-1, exercise = TRUE}

```

```{r position-adjustments-1-hint-1, eval = FALSE}
ggplot(mpg, aes(... = drv, fill = drv)) + 
  ...()
```

```{r position-adjustments-1-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = drv, fill = drv)) + 
  geom_bar()
```

```{r position-adjustments-1-check}
grade_this_code()
```
### 

This is a basic bar chart, but with the inside of each bar colored according to the value of `drv`. If instead we used `color` the border of each bar would have been colored.

### 

It may seem strange to use `drv` as the argument to both `x` and to `fill`. Yet this is a common approach. The two arguments are doing two different things but, in this case, both of those things are associated with `drv`. 

### Exercise 2

Change `fill = drv` to `fill = class`.


```{r position-adjustments-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r position-adjustments-2-hint-1, eval = FALSE}
ggplot(mpg, aes(x = drv, fill = ...)) + 
  geom_bar()
```

```{r position-adjustments-2-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = drv, fill = class)) + 
  geom_bar()
```

```{r position-adjustments-2-check}
grade_this_code()
```
### 

Note what happens if you map the fill aesthetic to another variable, like `class`: the bars are automatically stacked. Each colored rectangle represents a combination of `drv` and `class`.

### Exercise 3

The stacking is performed automatically using the position adjustment specified by the position argument. If you don’t want a stacked bar chart, you can use one of three other options: "identity", "dodge" or "fill". To explore, take the previous plot and add `alpha = 1/5, position = "identity"` to the call to `geom_bar()`.

```{r position-adjustments-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r position-adjustments-3-hint-1, eval = FALSE}
ggplot(mpg, aes(x = drv, fill = class)) + 
  geom_bar(alpha = ..., ... = "identity")
```

```{r position-adjustments-3-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = drv, fill = class)) + 
  geom_bar(alpha = 1/5, position = "identity")
```

```{r position-adjustments-3-check}
grade_this_code()
```

### 

`position = "identity"` will place each object exactly where it falls in the context of the graph. This is not very useful for bars, because it overlaps them. To see that overlapping we either need to make the bars slightly transparent by setting alpha to a small value.

### 

The identity position adjustment is more useful for 2d geoms, like points, where it is the default.

### Exercise 4

`position = "fill"` works like stacking, but makes each set of stacked bars the same height. Use the same code, but remove `fill = NA` and replace `"identity"` with `"fill"`.

```{r position-adjustments-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r position-adjustments-4-hint-1, eval = FALSE}

```

```{r position-adjustments-4-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = drv, fill = class)) + 
  geom_bar(position = "fill")
```

```{r position-adjustments-4-check}
grade_this_code()
```

### 

This makes it easier to compare proportions across groups. 

### Exercise 5

`position = "dodge"` places overlapping objects directly beside one another.  Replace `"fill"` with `"dodge"`.

```{r position-adjustments-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r position-adjustments-5-hint-1, eval = FALSE}
ggplot(mpg, aes(x = drv, fill = class)) + 
  geom_bar(position = ...)
```

```{r position-adjustments-5-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = drv, fill = class)) + 
  geom_bar(position = "dodge")
```

```{r position-adjustments-5-check}
grade_this_code()
```

### 

This makes it easier to compare individual values.

### Exercise 6

There’s one other type of adjustment that’s not useful for bar charts, but can be very useful for scatterplots. Recall our first scatterplot. Did you notice that the plot displays only 126 points, even though there are 234 observations in the dataset? Starting with our usual `ggplot(mpg, aes(x = displ, y = hwy))`, add `geom_point(position = "jitter")`

```{r position-adjustments-6, exercise = TRUE}

```

```{r position-adjustments-6-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point(... = "jitter")
```

```{r position-adjustments-6-solution, exercise.reveal_solution = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point(position = "jitter")
```

```{r position-adjustments-6-check}
grade_this_code()
```

### 

Adding randomness seems like a strange way to improve your plot, but while it makes your graph less accurate at small scales, it makes your graph more revealing at large scales. Because this is such a useful operation, **ggplot2** comes with a shorthand for `geom_point(position = "jitter")`: `geom_jitter()`.

To learn more about a position adjustment, look up the help page associated with each adjustment: `?position_dodge`, `?position_fill`, `?position_identity`, `?position_jitter`, and `?position_stack`.

## Coordinate Systems
### 

Coordinate systems are probably the most complicated part of **ggplot2**. The default coordinate system is the Cartesian coordinate system where the x and y positions act independently to determine the location of each point. 

### Exercise 1

Load the `maps` package with `library()`.

```{r coordinate-systems-1, exercise = TRUE}

```

```{r coordinate-systems-1-hint-1, eval = FALSE}
library(...)
```

```{r coordinate-systems-1-solution, exercise.reveal_solution = FALSE}
library(maps)
```

```{r coordinate-systems-1-check}
grade_this_code()
```

### 

The **maps** package includes the area of each matching region in the map is computed, and regions which match the same element of regions have their areas combined. Each region is assumed planar, with vertices specified by the x and y components of the map object.

### Exercise 2

Hit "Submit Code" to create an object with data for New Zealand and then create a plot.

```{r coordinate-systems-2, exercise = TRUE}
nz <- map_data("nz")

ggplot(nz, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black")

```

```{r coordinate-systems-2-hint-1, eval = FALSE}
nz <- map_data("nz")

ggplot(nz, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black")
```

```{r coordinate-systems-2-solution, exercise.reveal_solution = FALSE}
nz <- map_data("nz")

ggplot(nz, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black")
```

```{r coordinate-systems-2-check}
grade_this_code()
```

### 

We don’t have the space to discuss maps extensively in this tutorial but you can learn more in the [Maps](https://ggplot2-book.org/maps) chapter of [*ggplot2: Elegant Graphics for Data Analysis*](https://ggplot2-book.org/). 

### Exercise 3

`coord_quickmap()` sets the aspect ratio correctly for geographic maps. This is very important if you’re plotting spatial data with **ggplot2**. Add `coord_quickmap()` to the `ggplot()` call from the previous question.

```{r coordinate-systems-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-3-hint-1, eval = FALSE}
... +
  coord_quickmap()
```

```{r coordinate-systems-3-solution, exercise.reveal_solution = FALSE}
nz <- map_data("nz")

ggplot(nz, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black") +
  coord_quickmap()
```

```{r coordinate-systems-3-check}
grade_this_code()
```

### Exercise 4

`coord_polar()` uses polar coordinates. Polar coordinates reveal an interesting connection between a bar chart and a Coxcomb chart.

Consider this bar chart of the `diamonds` dataset that shows the count of each `clarity`.

```{r}
ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = clarity, fill = clarity), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1)
```

###

Add on `coord_polar()` to the end of the plot, don't forget to use the `+` operator.

```{r coordinate-systems-4, exercise = TRUE}
ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = clarity, fill = clarity), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1)
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-4-hint-1, eval = FALSE}
ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = clarity, fill = clarity), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  ...()
```

```{r coordinate-systems-4-solution, exercise.reveal_solution = FALSE}
ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = clarity, fill = clarity), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  coord_polar()
```

```{r coordinate-systems-4-check}
grade_this_code()
```

## The layered grammar of graphics
### 

Consider this overview:

````
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
````

Our new template takes seven parameters, the bracketed words that appear in the template. In practice, you rarely need to supply all seven parameters to make a graph because ggplot2 will provide useful defaults for everything except the data, the mappings, and the geom function.


### 

The seven parameters in the template compose the grammar of graphics, a formal system for building plots. The [grammar of graphics](https://vita.had.co.nz/papers/layered-grammar.pdf) is based on the insight that you can uniquely describe any plot as a combination of a dataset, a geom, a set of mappings, a stat, a position adjustment, a coordinate system, a faceting scheme, and a theme.


## Summary
### 

This tutorial covered [Chapter 9: Layers](https://r4ds.hadley.nz/layers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned how to use [**ggplot2**](https://ggplot2.tidyverse.org/) and related packages to create persuasive graphics. Key functions included [`geom_smooth()`](https://ggplot2.tidyverse.org/reference/geom_smooth.html),  [`geom_histogram()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html), and  [`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html).

Once you are finished:

-   Click the 'Download Grade' button below. This will download an html document of your grade summary.
-   Submit the downloaded html to your learning management system.

```{r}
grade_print_ui("grade")
```
