---
title: Workflow code style
tutorial:
  id: workflow-code-style
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
    css: ["css/nu-theme.css"]
runtime: shiny_prerendered
description: 'Tutorial for Chapter 4: Workflow code style'
---

<!-- Modification of r4ds.tutorials author: David Kane -->

```{r setup, include = FALSE}
library(learnr)
#library(tutorial.helpers)
library(tidyverse)
library(nycflights13)
library(knitr)
library(gradethis)
library(tutorialExtras)

gradethis_setup()

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local")
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r, context = "server"}
grade_server("grade")
```

```{r Name}
question_text(
  "Name:",
  answer_fn(function(value){
    if(length(value) >= 1 ) {
      return(mark_as(TRUE))
    }
    return(mark_as(FALSE) )
  }),
  correct = "submitted",
  allow_retry = FALSE )
```

```{r}
grade_button_ui(id = "grade")
```


## Introduction
### 

This tutorial covers [Chapter 4 Workflow: code style](https://r4ds.hadley.nz/workflow-style.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn the most important principles of code style including how to appropriately assign names and format code with proper spacing and line breaks.


## Names

The following exercises use the `flights` dataset from the `nycflights13` package.

###

### Exercise 1

Assign a short descriptive name to the following code that subsets the `flights` data frame to only include flights shorter than 60 minutes.

```{r names-1, exercise = TRUE}
___ <- flights |> filter(air_time < 60)
```

```{r names-1-hint-1, eval = FALSE}
___ <- flights |> filter(air_time < 60)
```

<!-- ```{r names-1-solution, exercise.reveal_solution = FALSE} -->
<!-- short_name <- flights |> filter(air_time < 60) -->
<!-- ``` -->

```{r names-1-check}

grade_this({
  
  user_ans <- str_split(.user_code, pattern = " <-")[[1]][1]
  allowed <- c("_", 0:9, letters)
  tmp <- sapply(strsplit(user_ans, ""), function(x){! (x %in% allowed)})

  if (sum(tmp) == 0) {
    pass("Great work!")
  }
  fail()
  
})
```


## Spaces

For the following exercises, correct the code to have appropriate spacing.

###

### Exercise 1

<!-- ```{r, eval=FALSE, echo=TRUE} -->
<!-- z<-(a+b)^2/d -->
<!-- ``` -->

<!-- ```{r spaces-1} -->
<!-- question_text( -->
<!--   "", -->
<!--   answer("z <- (a + b)^2 / d", correct = TRUE), -->
<!--   allow_retry = TRUE -->
<!--   ) -->
<!-- ``` -->

```{r spaces-1-setup}
a <- 2
b <- 3
d <- 4
```

```{r spaces-1, exercise = TRUE}
z<-(a+b)^2/d
```

```{r spaces-1-solution, exercise.reveal_solution = FALSE}
z <- (a + b)^2 / d
```

```{r spaces-1-check}
grade_this({
  # custom checking code appears here
  if (identical(.user_code, .solution_code)) {
    pass("Great work!")
  }
  fail("Try again!")
})

```

### Exercise 2

```{r spaces-2-setup}
x <- rnorm(20, 100, sqrt(10))
```

```{r spaces-2, exercise = TRUE, results='hide'}
mean (x ,na.rm=TRUE)
```

```{r spaces-2-solution, exercise.reveal_solution = FALSE}
mean(x, na.rm = TRUE)
```

```{r spaces-2-check}
grade_this({
  # custom checking code appears here
  if (identical(.user_code, .solution_code)) {
    pass("Great work!")
  }
  fail("Try again!")
})
```


## Pipes

`|>` should always have a space before it and should typically be the last thing on a line.

###

For the following exercises, correct the code to have appropriate pipe formatting.

### Exercise 1

```{r pipes-1, exercise = TRUE, results='hide'}
flights|>filter(!is.na(arr_delay), !is.na(tailnum))|>count(dest)
```


```{r pipes-1-solution, exercise.reveal_solution = FALSE}
flights |> 
  filter(!is.na(arr_delay), !is.na(tailnum)) |> 
  count(dest)
```

```{r pipes-1-check}
grade_this_code(
  correct = glue::glue(
  "Complete!", 
  "Verify that it matches the solution, provided once you click `Continue` below. The auto-grader has problems checking syntax so we are giving completion credit. Check that your solution is correct!", .sep = "<br><br>")
)
```

### Exercise 1 Solution

```{r, eval=FALSE, echo=TRUE}
flights |> 
  filter(!is.na(arr_delay), !is.na(tailnum)) |> 
  count(dest)
```


### Exercise 2

```{r pipes-2, exercise = TRUE, results='hide'}
flights |>
  group_by(
    tailnum
  ) |> 
  summarize(delay = mean(arr_delay, na.rm = TRUE), n = n())
```


```{r pipes-2-solution, exercise.reveal_solution = FALSE}
flights |>  
  group_by(tailnum) |> 
  summarize(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )
```

```{r pipes-2-check}
grade_this_code(
  correct = glue::glue(
  "Complete!", 
  "Verify that it matches the solution, provided once you click `Continue` below. The auto-grader has problems checking syntax so we are giving completion credit. Check that your solution is correct!", .sep = "<br><br>")
)
```

### Exercise 2 Solution

```{r, eval=FALSE, echo=TRUE}
flights |>  
  group_by(tailnum) |> 
  summarize(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )
```


## ggplot2

The `+` operator follows the same formatting as the pipe `|>`

###

### Exercise 1

Correct the code to have appropriate formatting.

```{r ggplot2-1, exercise = TRUE, results='hide', warning=FALSE, message=FALSE}
speed_v_distance_plot <- flights |> 
  group_by(dest) |> 
  summarize(
    distance = mean(distance),
    speed = mean(distance / air_time, na.rm = TRUE)
  ) |> 
  ggplot(aes(x = distance, y = speed)) +
  geom_smooth(
    method = "loess", span = 0.5, se = FALSE, 
    color = "white", linewidth = 4
  ) + geom_point()
```


```{r ggplot2-1-solution, exercise.reveal_solution = FALSE}
speed_v_distance_plot <- flights |> 
  group_by(dest) |> 
  summarize(
    distance = mean(distance),
    speed = mean(distance / air_time, na.rm = TRUE)
  ) |> 
  ggplot(aes(x = distance, y = speed)) +
  geom_smooth(
    method = "loess",
    span = 0.5,
    se = FALSE, 
    color = "white", 
    linewidth = 4
  ) +
  geom_point()
```

```{r ggplot2-1-check}
grade_this_code(
  correct = glue::glue(
  "Complete!", 
  "Verify that it matches the solution, provided once you click `Continue` below. The auto-grader has problems checking syntax so we are giving completion credit. Check that your solution is correct!", .sep = "<br><br>")
)
```

### Exercise 1 Solution

```{r, eval=FALSE, echo=TRUE}
speed_v_distance_plot <- flights |> 
  group_by(dest) |> 
  summarize(
    distance = mean(distance),
    speed = mean(distance / air_time, na.rm = TRUE)
  ) |> 
  ggplot(aes(x = distance, y = speed)) +
  geom_smooth(
    method = "loess",
    span = 0.5,
    se = FALSE, 
    color = "white", 
    linewidth = 4
  ) +
  geom_point()
```

## Sectioning comments

```{r sec-comment-1-q}
quiz(
  caption = "",
  question("What is the 'shortcut' for creating sectioning comments?",
       answer("ctrl/cmd + shift + R", correct = TRUE),
       answer("ctrl/cmd + shift + C"),
       answer("alt + -"),
       answer("shift + # + tab"),
       answer("alt + #"),
       answer("ctrl/cmd + shift + #"),
       random_answer_order = TRUE,
       allow_retry = TRUE),
  question_text("What package provides a quick and easy way to improve the quality of poorly styled code?",
       answer("styler", correct = TRUE),
       allow_retry = TRUE)
)
```


## Summary

### 

This tutorial covered [Chapter 4 Workflow: code style](https://r4ds.hadley.nz/workflow-style.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned the most important principles of code style including how to appropriately assign names and format code with proper spacing and line breaks.

Once you are finished:

-   Click the 'Download Grade' button below. This will download an html document of your grade summary.
-   Submit the downloaded html to your learning management system.

```{r}
grade_print_ui("grade")
```
